# Missing values

```{r}
# Import necessary packages.
library(ggplot2)
library(tidyverse)
library(patchwork)
```

```{r}
# Create a function which can be used to analyze our data.
plot_missing <- function(data, percent = TRUE){
  
  # Find numbers of columns and rows of the dataset
  num_rows <- nrow(data)
  num_cols <- ncol(data)
  
  # Find and count the missing patterns
  missing_patterns <- data.frame(is.na(data)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()
  
  # Modify missing patterns based on the percent option
  if (percent == TRUE){
  # Modify missing patterns to generate main graph
  missing_patterns_main <- missing_patterns %>%
    mutate(id = row_number(), .before = 1) %>%
    pivot_longer(cols = -c("id", "count"), 
                 names_to = "variables", 
                 values_to = "value") %>%
    group_by(variables) %>%
    mutate(var_count = 100 * sum(count[value == TRUE])/num_rows) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(complete = as.character(ifelse(sum(value) == 0, 1, 0))) %>%
    ungroup()
  
  # Modify missing patterns to generate top graph
  missing_patterns_top <- missing_patterns_main %>%
    group_by(variables) %>%
    summarise(missing_rows = 100 * sum(count[value == TRUE])/num_rows) %>%
    ungroup()
  
  # Modify missing patterns to generate right graph
  missing_patterns_right <- missing_patterns %>%
    mutate(id = row_number(), .before = 1) %>%
    mutate(count = 100 * count/num_rows, 
           complete = as.character(ifelse(rowSums(.[2:(num_cols + 1)]) == 0, 1, 0))) %>%
    select(id, count, complete)
  }
  
  else if (percent == FALSE){
  # Modify missing patterns to generate main graph
  missing_patterns_main <- missing_patterns %>%
    mutate(id = row_number(), .before = 1) %>%
    pivot_longer(cols = -c("id", "count"), 
                 names_to = "variables", 
                 values_to = "value") %>%
    group_by(variables) %>%
    mutate(var_count = sum(count[value == TRUE])) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(complete = as.character(ifelse(sum(value) == 0, 1, 0))) %>%
    ungroup()
  
  # Modify missing patterns to generate top graph
  missing_patterns_top <- missing_patterns_main %>%
    group_by(variables) %>%
    summarise(missing_rows = sum(count[value == TRUE])) %>%
    ungroup()
  
  # Modify missing patterns to generate right graph
  missing_patterns_right <- missing_patterns %>%
    mutate(id = row_number(), .before = 1) %>%
    mutate(complete = as.character(ifelse(rowSums(.[2:(num_cols + 1)]) == 0, 1, 0))) %>%
    select(id, count, complete)
  }
  
  else{
    print("Wrong data input or inproper argument. Please try again. ")
  }

  # Set colors for main graph (not using defaults)
  tile_colors <- c("grey60", "mediumorchid4")

  # Plot main graph
  main_plot <- ggplot(missing_patterns_main, 
                      aes(x = reorder(variables, -var_count), 
                          y = fct_rev(factor(id)), 
                          fill = value, 
                          alpha = complete)) + 
          geom_tile(color = "white", show.legend = FALSE) + 
          scale_fill_manual(values = tile_colors) + 
          scale_alpha_manual(values = c("1" = 1, "0" = 0.5)) + 
          labs(x = "variables", y = "missing pattern") + 
          theme_bw() + 
          theme(panel.grid.major = element_blank(),
                panel.grid.minor = element_blank())
  
  # Plot top graph
  top_plot <- ggplot(missing_patterns_top) + 
          geom_col(aes(x = reorder(variables, -missing_rows),
                       y = missing_rows), 
                   fill = "cornflowerblue") + 
          labs(title = "Missing value patterns", 
               x = "") + 
          theme_bw() + 
          theme(panel.grid.major.x = element_blank(), 
                panel.grid.minor.x = element_blank())
  
  # Add labels and title of top graph based on the percent option
  if (percent == TRUE){
    top_plot <- top_plot + 
      scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25))+
      ylab("% rows missing: ")
  }
  else{
    top_plot <- top_plot + 
      ylab("num rows missing: ")
  }

  # Plot right graph
  right_plot <- ggplot(missing_patterns_right, 
                       aes(x = count, 
                           y = fct_rev(factor(id)), 
                           fill = complete)) + 
          geom_col() + 
          scale_fill_manual(values = c("0" = "cornflowerblue", "1" = "blue")) + 
          ylab("") + 
          theme_bw() + 
          theme(panel.grid.major.y = element_blank(), 
                panel.grid.minor.y = element_blank(), 
                legend.position = "none")
  
  # Add labels and title of right graph based on the percent option
  if (percent == TRUE){
    right_plot <- right_plot + 
      scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25)) +
      xlab("% rows")
  }
  else{
    right_plot <- right_plot + 
      xlab("row count")
  }
  
  # Find position to label complete cases
  complete_cases <- unique(missing_patterns_main$id[which(missing_patterns_main$complete == "1")])
  x = as.integer(ncol(data)/2) + 1
  for (i in complete_cases){
    y = nrow(missing_patterns) - as.integer(i) + 1
    main_plot <- main_plot + 
      geom_text(aes(x, y, label = "complete case"), 
                size = 10, family = "Times New Roman", 
                hjust = 0.6, show_guide = FALSE)
}

  # Set layout for plots assembly
  layout <- c(area(1, 1, 1, 4), area(2, 1, 5, 4), area(2, 5, 5, 5))

  # Assemble graphs
  top_plot + main_plot + right_plot + plot_layout(design = layout)
}
```


```{r}
# Load data
df <- read.csv("species.csv", sep = ",", encoding = "UTF-8")
```

```{r}
# Shorten the column names so they don't overlap in the plot.
df_colnames <- colnames(df)
colnames(df) <- c('SP','ST','CNT','LA','LO','COUN','COPRE','ELE','ELEA','DP','DPA','M','Y','SK')
```

Visualize the missing pattern of the data which we have purged before.
```{r}
plot_missing(df)
```

Observing the plot, we can easily find out some correlations among some factors. LA means latitude and LO means longitude, so if one of them is missing, the other one must be missing too. The plot reveals this correlation. The similar correlation also appears between DP(depth) and DPA(depth accuracy), and EA(elevation accuracy) and ELE(elevation). Another correlation is between Y(year) and M(month). Year is more easily be recorded than month.

DP, DPA, COPRE(coordinate precision), CNT are more likely to be missing. We might use elevation to conduct some results since nearly half of them are not missing. COUN(coordinate uncertainty in meter) has no missing because we first downloaded data with limitation of COUN smaller than 10000. And SP(species), SK(species key) has no missing data because at chapter 3 we purge all missing species key data. ST(state), Y has no missing at all. ST(state) and Y(year) is much easier to record. 

Although LA, LO and M have some missing values, these cases are actually very small. We can notice the top graph that it is not very obvious to show that these columns have missing values. As for the right graph, we can find that the missing pattern that these columns have are smallest subset in the whole data. Therefore, we can think these columns have no missing values.
