# Results
## The correlation between latitude and year.
```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
# load data
df <- read.csv("species.csv", sep = ",", encoding = "UTF-8")
colnames(df) <- c('SP','ST','CNT','LA','LO','COUN','COPRE','ELE','ELEA','DP','DPA','M','Y','SK')
```

In order to find out the correlation between latitude and year, we need to purge some data whose latitude and year may not be relative. For instance, this is one of species called Utricularia dichotoma, the correlation coefficient between latitude and year is very small. 

```{r}
la_sp <- subset(df, SP == "Utricularia dichotoma", select = c('SP','LA','Y'))
la_sp <- aggregate(la_sp$LA, list(la_sp$Y), FUN=mean)
cor.test(la_sp$Group.1, la_sp$x, method = "spearman")$estimate[[1]]
```

The scatter plot of its latitude and year seems messy. It is hard to describe the correlation between them. The reason why it happens is probably wrong records or the characteristic of plants. Utricularia dichotoma usually grows in wet locations and the moisture must be more important than temperature to the living of the species causing the mess of the scatter plot. Therefore, if the correlation coefficient is smaller than 0.3, we think no correlation exists.

```{r}
ggplot(la_sp, aes(Group.1, x)) +
  geom_point() + 
  geom_smooth(method = "lm") +
  ylab("average latitude") +
  xlab("year") +
  ggtitle("Utricularia dichotoma: latitude ~ year")
```

We calculate the correlation coefficient of latitude and year of each species, and choose its absolute value which is not less than 0.3. There are 19 species met the condition. Here are the species, correlation coefficients and slopes.

```{r}
# 
sp_names <- unique(df$SP)
sp_use <- NULL
corr_list <- NULL
slope <- NULL

all_sp_use <- NULL
all_corr_list <- NULL

for (sp in sp_names) {
  la_sp <- subset(df, SP == sp, select = c('SP','LA','Y'))
  la_sp <- aggregate(la_sp$LA, list(la_sp$Y), FUN = mean)
  corr <- cor.test(la_sp$Group.1, la_sp$x, method = "pearson")$estimate
  if (abs(corr) >= 0.3) {
    sp_use <- append(sp_use, sp)
    corr_list <- append(corr_list, corr)
    slope <- append(slope, lm(la_sp$x ~ la_sp$Group.1)[[1]][[2]])
  }
  all_sp_use <- append(all_sp_use, sp)
  all_corr_list <- append(all_corr_list, corr)
}
la_y_corr <- data.frame('species' = sp_use, 'corr' = corr_list, 'slope' = slope)
all_la_y_corr <- data.frame('species' = all_sp_use, 'corr' = all_corr_list)
la_y_corr
```

Observing the table, we can find that there are 3 species have positive correlation coefficients and 16 species have negative ones. It is worth noting that all latitudes are smaller than 0 since Australia is located in the southern hemisphere. Therefore, the negative correlation means the latitude gets higher as years pass. We show the correlation between latitude and years among 63 species as follows. The result is that about 70% of the species have no correlation between latitude and year. Nearly 25% of the species have negative correlations and 5% of the species have positive ones.

```{r}
positive_species <- subset(la_y_corr, slope > 0, select = c(species, corr))
negative_species <- subset(la_y_corr, slope < 0, select = c(species, corr))
relation <- c('positive', 'negative', 'norelation')
count <- c(nrow(positive_species), 
           nrow(negative_species), 
           length(sp_names) - nrow(negative_species) - nrow(positive_species))

pie <- data.frame('relation'= relation, 'count' = count)
mylabel <- as.vector(pie$relation)
mylabel <- paste(mylabel, "(", round(pie$count / sum(pie$count) * 100, 2), "%) ", sep = "") 

ggplot(pie, aes(x = "", y = count, fill = relation)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  ggtitle("Correlation Between Latitude and Year") +
  labs(x = "", y = "") +
  scale_fill_discrete(breaks = pie$relation, labels = mylabel)
```

We are interested in whether the changes of latitude are significant. The null hypothesis is that the correlation subjects to the normal distribution, which means that the change of is normal. Q-Q plot, density curve and Shapiro Wilk test result are presented. We can find that since p-value is very large, it does subject to the normal distribution, which means that the change of latitude is normal. But we still find that the density curve leans to left a little bit, which means the latitude has a tendency to get higher.

```{r}
shapiro.test(all_la_y_corr$corr)
qqnorm(all_la_y_corr$corr)
qqline(all_la_y_corr$corr, col = "red")

ggplot(all_la_y_corr, aes(corr)) +
  geom_histogram(aes(y = ..density..),
                 fill = "lightblue",
                 color = "black",
                 bins = 15) +
  geom_density(lwd=0.5) +
  stat_function(fun = dnorm, args = list(mean = mean(all_la_y_corr$corr), sd = sd(all_la_y_corr$corr)), color = "red", lwd = 1.5)
```

We are also interested in different correlations in different class: Liliopsida and Magnoliopsida. We can see that the latitudes of Magnoliosida plants are more likely get higher, but the correlation is not significant. And Liliopsida plants do not have any correlation between latitudes and years.

```{r}
test2 <- unique(subset(data, species %in% all_sp_use, select = c(class, species)))
test2 <- subset(left_join(test2, all_la_y_corr, by="species"), class != "Pinopsida")

ggplot(test2, aes(corr, reorder(class, -corr, median))) +
  geom_boxplot() +
  ylab("class") +
  xlab("correlation") +
  ggtitle("Correlation Between Latitude and Year in Different Class") +
  theme(plot.title = element_text(size = 15))
```

## The correlation between elevation and year.

We do the same thing to analyze the correlation between elevation and year. The result is that about 62% of the species have no correlation. Around 25% of the species have positive correlations and 13% of the species have negative ones.

```{r}
sp_names <- unique(df$SP)
sp_use <- NULL
corr_list <- NULL
slope <- NULL

all_sp_use <- NULL
all_corr_list <- NULL

for (sp in sp_names) {
  ele_sp <- subset(df, SP == sp, select = c('SP','ELE','Y'))
  ele_sp <- ele_sp %>%
            filter(!is.na(ELE))
  ele_sp <- aggregate(ele_sp$ELE, list(ele_sp$Y), FUN = mean)
  corr <- cor.test(ele_sp$Group.1, ele_sp$x, method = "pearson")$estimate
  if (abs(corr) >= 0.3) {
    sp_use <- append(sp_use, sp)
    corr_list <- append(corr_list, corr)
    slope <- append(slope, lm(ele_sp$x ~ ele_sp$Group.1)[[1]][[2]])
  }
  all_sp_use <- append(all_sp_use, sp)
  all_corr_list <- append(all_corr_list, corr)  
}
ele_y_corr <- data.frame('species' = sp_use, 'corr' = corr_list, 'slope' = slope)
all_ele_y_corr <- data.frame('species' = all_sp_use, 'corr' = all_corr_list)

positive_species <- subset(ele_y_corr, slope > 0, select = c(species, corr))
negative_species <- subset(ele_y_corr, slope < 0, select = c(species, corr))
relation <- c('positive', 'negative', 'norelation')
count <- c(nrow(positive_species), 
           nrow(negative_species), 
           length(sp_names) - nrow(negative_species) - nrow(positive_species))

pie <- data.frame('relation'= relation, 'count' = count)
mylabel <- as.vector(pie$relation)
mylabel <- paste(mylabel, "(", round(pie$count / sum(pie$count) * 100, 2), "%) ", sep = "") 

ggplot(pie, aes(x = "", y = count, fill = relation)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  ggtitle("Correlation Between Elevation and Year") +
  labs(x = "", y = "") +
  scale_fill_discrete(breaks = pie$relation, labels = mylabel)
```

We also present the Q-Q plot, density curve and Shapiro Wilk test result to show whether the correlation subject to normal distribution. At 10 percent level of significance, we can reject the null hypothesis. It means that the change of elevation is not normal in the last 50 years. We can notice that the density curve leans to left, which means that the elevation has a tendency to get higher.

```{r}
shapiro.test(all_ele_y_corr$corr)
qqnorm(all_ele_y_corr$corr)
qqline(all_ele_y_corr$corr, col = "red")

ggplot(all_ele_y_corr, aes(corr)) +
  geom_histogram(aes(y = ..density..),
                 fill = "lightblue",
                 color = "black",
                 bins = 20) +
  geom_density(lwd=0.5) +
  stat_function(fun = dnorm, args = list(mean = mean(all_ele_y_corr$corr), sd = sd(all_ele_y_corr$corr)), color = "red", lwd = 1.5)
```

Different correlations in different class are also presented as follows. The elevations of Liliopsida plants are more likely get higher as years pass, but the correlation is not significant. It seems the Magnoliopsida do not have such correlation since its center lies near zero.

```{r}
test2 <- unique(subset(data, species %in% all_sp_use, select = c(class, species)))
test2 <- subset(left_join(test2, all_ele_y_corr, by="species"), class != "Pinopsida")

ggplot(test2, aes(corr, reorder(class, -corr, median))) +
  geom_boxplot() +
  ylab("class") +
  xlab("correlation") +
  ggtitle("Correlation Between Elevation and Year in Different Class") +
  theme(plot.title = element_text(size = 15))
```

Result:

1. The latitude of occurrences of Australia plants in the last 50 years has a tendency to get higher, but the change is minor.

2. The elevation of occurrences of Australia plants in the last 50 years has a tendency to get higher and the change is significant to some extent.